---
title: "New Look Q2"
author: "Colin Dassow and Chelsey Neiman"
date: "10/30/2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(ggplot2)
library(deSolve)
library(ggpubr)
library(metR)
source('C:/Users/jones/BoxSync/NDstuff/Dissertation/4/MultiSppModel/q2Func.R')

```


```{r, echo=F,cache=T, fig.cap='Figure 1. Model run to equilibrium over a range of harvest parameters for species 1, species 2 harvest is held constant at 1.8. No stocking. Top panel shows equilibrium abundances for the range of harvests when species 1 is initially dominant. Bottom panel shows equilibrium abundances for the range of harvests when species 2 is initially dominant.'}
#other option for figure 1

store=data.frame(qEs=seq(0,8,length.out=30),A1=0,A2=0,J1=0,J2=0)
y0=c(100,10,0,0)
tstep=1:300
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(s1=0.1,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.002,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.001,cJ2A1=0.5,cJ2J1=0.002,v2=1)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
store2=data.frame(qEs=seq(0,8,length.out=30),A1=0,A2=0,J1=0,J2=0)
y0=c(10,100,0,0)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(s1=0.1,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.002,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.001,cJ2A1=0.5,cJ2J1=0.002,v2=1)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}

par(mfcol=c(2,1), mar=c(1,1,3,1), oma=c(4,4,1,1))
plot(store$qEs,store$A1,lwd=3,type='l',ylim=c(0,max(store[,2:3],na.rm = T)),ylab = "",xlab = "", main = "Sp1 > Sp2")
lines(store$qEs,store$A2,lwd=3,col='grey')
legend("topright",legend = c("sp 1", "sp 2"), lty=1, lwd=2, col = c("black","grey"),bty="n")
plot(store2$qEs,store2$A1,lwd=3, type='l', ylim = c(0, max(store2[,2:3],na.rm = T)),ylab = "", xlab = "", main = "Sp2 > Sp1")
lines(store2$qEs,store2$A2,lwd=3,col='grey')
legend("topright",legend = c("sp 1", "sp 2"), lty=1, lwd=2, col = c("black","grey"),bty="n")
mtext("Abundance", side = 2, outer = T, line = 2.5)
mtext("Species 1 Harvest Rate (q*E)", side = 1, outer = T, line = 2.5)
```


```{r, echo=FALSE, cache=T, fig.dim=c(7,7), fig.cap="Figure 2. Effect of 'multispecies' thinking on either maintaining sp1 dominance (A&C) or flipping to a state where sp 1 dominates (B&D). This is described for systems where a manager managers only the focoal species  (A&B) or both species (B&D). Green = sp1 >sp2, red = sp2 > sp1. Blue line marks boundary where sp1 > sp2 by at least 100 individuals. This is shown as an exmaple of what a manager may care about, not just that sp1 is a little more abundant than sp2 but that the diffferent meets some minimum requirement."}
##### HEATMAPS #####

##### HEATMAPS_NEW 9.28.2020 #####

h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))

##### MAINTAIN W/O sp2 harv ####
#matrix to hold output, starting with different harvest levels on each species

qEs=seq(0,8,length.out = 30)
sto=seq(0,2000, length.out = 30)
df=expand.grid(X=qEs,Y=sto)
df$A1=numeric(nrow(df))
df$A2=numeric(nrow(df))
df$J1=numeric(nrow(df))
df$J2=numeric(nrow(df))
minDiff=100

for(i in 1:nrow(df)){
  tstep=1:100
  qE1Fun=approxfun(x=tstep,y=rep(df$X[i], length(tstep)))
  qE2Fun=approxfun(x=tstep,y=rep(0, length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(df$Y[i],length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
  y0=c(1000,100,0,0)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  df$A1[i]=sim[nrow(sim)-1,2]
  df$A2[i]=sim[nrow(sim)-1,3]
  df$J1[i]=sim[nrow(sim)-1,4]
  df$J2[i]=sim[nrow(sim)-1,5]
}

df$diff=df$A1-df$A2
df$outcome=ifelse(df$A1-df$A2 < minDiff,"darkgreen", "darkred")
vz=ggplot(data = df)+theme_classic()
a=vz+geom_tile(aes(x=df$X, y=df$Y, fill=df$diff))+
  scale_fill_gradient2(low="darkred", high="darkgreen", mid="white", midpoint=1, name="Sp1 Dominance", breaks=c(min(df$diff), 1, max(df$dif)), labels=c("sp2", "even", "sp1"))+
  labs(x="Harvest Rate", y="Stocked fish", title = "Manage species 1 only")+
  geom_contour(aes(x=df$X, y=df$Y, z=df$diff), breaks = c(minDiff))

#### FLIP W/O sp2 harv ####

qEs=seq(0,8,length.out = 30)
sto=seq(0,2000, length.out = 30)
df2=expand.grid(X=qEs,Y=sto)
df2$A1=numeric(nrow(df2))
df2$A2=numeric(nrow(df2))
df2$J1=numeric(nrow(df2))
df2$J2=numeric(nrow(df2))


for(i in 1:nrow(df)){
  tstep=1:100
  qE1Fun=approxfun(x=tstep,y=rep(df2$X[i], length(tstep)))
  qE2Fun=approxfun(x=tstep,y=rep(0, length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(df2$Y[i],length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
  y0=c(100,1000,0,0)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  df2$A1[i]=sim[nrow(sim)-1,2]
  df2$A2[i]=sim[nrow(sim)-1,3]
  df2$J1[i]=sim[nrow(sim)-1,4]
  df2$J2[i]=sim[nrow(sim)-1,5]
}

df2$diff=df2$A1-df2$A2
df2$outcome=ifelse(df2$A1-df2$A2 < minDiff,"darkgreen", "darkred")
vz=ggplot(data = df2)+theme_classic()
b=vz+geom_tile(aes(x=df2$X, y=df2$Y, fill=df2$diff))+
  scale_fill_gradient2(low="darkred", high="darkgreen", mid="white", midpoint=1, name="Sp1 Dominance", breaks=c(min(df2$diff), 1, max(df2$dif)), labels=c("sp2", "even", "sp1"))+
  labs(x="Harvest Rate", y="Stocked fish", title = "Manage species 1 only")+
  geom_contour(aes(x=df2$X, y=df2$Y, z=df2$diff), breaks = c(minDiff))

##### MAINTAIN W/ sp2 harv ####
#matrix to hold output, starting with different harvest levels on each species

qEs=seq(0,8,length.out = 30)
sto=seq(0,2000, length.out = 30)
dfwo=expand.grid(X=qEs,Y=sto)
dfwo$A1=numeric(nrow(dfwo))
dfwo$A2=numeric(nrow(dfwo))
dfwo$J1=numeric(nrow(dfwo))
dfwo$J2=numeric(nrow(dfwo))

for(i in 1:nrow(dfwo)){
  tstep=1:100
  qE1Fun=approxfun(x=tstep,y=rep(dfwo$X[i], length(tstep)))
  qE2Fun=approxfun(x=tstep,y=rep(2, length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(dfwo$Y[i],length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
  y0=c(1000,100,0,0)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  dfwo$A1[i]=sim[nrow(sim)-1,2]
  dfwo$A2[i]=sim[nrow(sim)-1,3]
  dfwo$J1[i]=sim[nrow(sim)-1,4]
  dfwo$J2[i]=sim[nrow(sim)-1,5]
}

dfwo$diff=dfwo$A1-dfwo$A2
dfwo$outcome=ifelse(dfwo$A1-dfwo$A2 < minDiff,"darkgreen", "darkred")
vz=ggplot(data = dfwo)+theme_classic()
c=vz+geom_tile(aes(x=dfwo$X, y=dfwo$Y, fill=dfwo$diff))+
  scale_fill_gradient2(low="darkred", high="darkgreen", mid="white", midpoint=1, name="Sp1 Dominance", breaks=c(min(dfwo$diff), 1, max(dfwo$dif)), labels=c("sp2", "even", "sp1"))+
  labs(x="Harvest Rate", y="Stocked fish", title = "manage both, Sp2 CPUE=2")+
  geom_contour(aes(x=dfwo$X, y=dfwo$Y, z=dfwo$diff), breaks = c(minDiff))

#### FLIP W/ sp2 harv ####

qEs=seq(0,8,length.out = 30)
sto=seq(0,2000, length.out = 30)
dfwo2=expand.grid(X=qEs,Y=sto)
dfwo2$A1=numeric(nrow(dfwo2))
dfwo2$A2=numeric(nrow(dfwo2))
dfwo2$J1=numeric(nrow(dfwo2))
dfwo2$J2=numeric(nrow(dfwo2))

for(i in 1:nrow(dfwo2)){
  tstep=1:100
  qE1Fun=approxfun(x=tstep,y=rep(dfwo2$X[i], length(tstep)))
  qE2Fun=approxfun(x=tstep,y=rep(2, length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(dfwo2$Y[i],length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
  y0=c(100,1000,0,0)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  dfwo2$A1[i]=sim[nrow(sim)-1,2]
  dfwo2$A2[i]=sim[nrow(sim)-1,3]
  dfwo2$J1[i]=sim[nrow(sim)-1,4]
  dfwo2$J2[i]=sim[nrow(sim)-1,5]
}

dfwo2$diff=dfwo2$A1-dfwo2$A2
dfwo2$outcome=ifelse(dfwo2$A1-dfwo2$A2 < minDiff,"darkgreen", "darkred")
vz=ggplot(data = dfwo2)+theme_classic()
d=vz+geom_tile(aes(x=dfwo2$X, y=dfwo2$Y, fill=dfwo2$diff))+
  scale_fill_gradient2(low="darkred", high="darkgreen", mid="white", midpoint=1, name="Sp1 Dominance", breaks=c(min(dfwo2$diff), 1, max(dfwo2$dif)), labels=c("sp2", "even", "sp1"))+
  labs(x="Harvest Rate", y="Stocked fish", title = " Manage both, Sp2 CPUE=2")+
  geom_contour(aes(x=dfwo2$X, y=dfwo2$Y, z=dfwo2$diff), breaks = c(minDiff))

#comparing with and without hysteresis plots together using cowplot package
#dev.new()
ggarrange(a,b,c,d,labels = c("A","B","C","D"), nrow = 2,ncol = 2, common.legend = T, legend = 'bottom')

```


```{r, echo=F, fig.cap="Figure 3. Differences between species 1 and 2 at equillibrium. Here managers can stock species 1 (y axis) or increase harvest on it's predator (species 2, x axis).  I was trying to show that adding harvest of sp2 in makes it much easier to reach management goals. I held sp1 harvest constant at 4. If you look at panel B from the previous figure you see that with no amount of stocking at a harvest of 4 should we be able to flip a system from sp2 to sp1 dominant. Here we see that if you add some harvest of sp2 in addition to stocking then you can stock less or stock just as much but tip the scales farther towards sp1. Not sure if this is the best way to vizualize this but it's a start."}

#plot to look at the cost/benefits of stocking or predator reduction for managing a focal species. 

qEs=seq(0,8,length.out = 30)
sto=seq(0,2000, length.out = 30)
dfT=expand.grid(X=qEs,Y=sto)
dfT$A1=numeric(nrow(dfT))
dfT$A2=numeric(nrow(dfT))
dfT$J1=numeric(nrow(dfT))
dfT$J2=numeric(nrow(dfT))

for(i in 1:nrow(dfT)){
  tstep=1:100
  qE1Fun=approxfun(x=tstep,y=rep(4, length(tstep)))
  qE2Fun=approxfun(x=tstep,y=rep(dfT$X[i], length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(dfT$Y[i],length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
      s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
  y0=c(100,1000,0,0)
  sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)
  dfT$A1[i]=sim[nrow(sim)-1,2]
  dfT$A2[i]=sim[nrow(sim)-1,3]
  dfT$J1[i]=sim[nrow(sim)-1,4]
  dfT$J2[i]=sim[nrow(sim)-1,5]
}

dfT$diff=dfT$A1-dfT$A2
dfT$qNorm=(dfT$X-min(dfT$X))/(max(dfT$X)-min(dfT$X))
dfT$sNorm=(dfT$Y-min(dfT$Y))/(max(dfT$Y)-min(dfT$Y))
#dfT$outcome=ifelse(dfT$A1-dfT$A2 < minDiff,"darkgreen", "darkred")
vz=ggplot(data = dfT)+theme_classic()
vz+geom_tile(aes(x=dfT$qNorm, y=dfT$sNorm, fill=dfT$diff))+
  scale_fill_gradient2(low="darkred", high="darkgreen", mid="white", midpoint=1, name="Sp1 \nDominance", breaks=c(min(dfT$diff), 1, max(dfT$diff)), labels=c("sp2", "even", "sp1"))+
  labs(x="Sp2 harvest Rate", y="Sp1 stocked fish")#+
  #geom_contour(aes(x=dfT$X, y=dfT$Y, z=dfT$diff))+
  #geom_label_contour(aes(x=dfT$X, y=dfT$Y, z=dfT$diff))
```

```{r, echo=F, fig.align="center", fig.cap="Figure 4. Delaying a transitions. Slow moving variable 'h'represents changing habitat availability which will inevitably flip system from sp1 dominated to sp2 (panel A). The flip in system state can be delayed through either stocking of the desired species (panel B), harvest of it's competitior (panel C), or perhaps prevented altogether by stocking and harvesting (panel D)."}

#figure on delaying a transition

#running to sp1 dominating before habitat decline
tstep=1:200
h1Fun=approxfun(x=tstep,y=c(rep(10,200)))
h2Fun=approxfun(x=tstep,y=c(rep(10,200)))
qE1Fun=approxfun(x=tstep,y=c(rep(0,length(tstep))))
qE2Fun=approxfun(x=tstep,y=c(rep(0,length(tstep))))
st1Fun=approxfun(x=tstep,y=c(rep(0,length(tstep))))
st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
    s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
y0=c(500,100,0,0)
simPre=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)

#habitat decline and stocking
tstep=1:500
h1Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,0, length.out = 100), rep(0,300)))
h2Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,10, length.out = 100),rep(10,300)))
qE1Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,4,length.out = 200), rep(4,200)))
qE2Fun=approxfun(x=tstep,y=c(rep(0,500)))
st1Fun=approxfun(x=tstep,y=c(rep(0,500)))
st2Fun=approxfun(x=tstep,y=c(rep(4,500)))#this is to simulate increase in fecundity, could change the model to make f time dependent, or have some amount of harvest that declines to 0
p=c(s1=0.1,m1=0.5,cJ1A1=0.002,cJ1A2=0.5,cJ1J2=0.003,v1=1,
    s2=0.1,m2=0.5,cJ2A2=0.002,cJ2A1=0.3,cJ2J1=0.003,v2=1)
y0=simPre[199,2:5]
sim=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)

#stocking delays the transition to sp2
tstep=1:500
h1Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,0, length.out = 100), rep(0,300)))
h2Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,10, length.out = 100),rep(10,300)))
qE1Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,4,length.out = 200), rep(4,200)))
qE2Fun=approxfun(x=tstep,y=c(rep(0,500)))
st1Fun=approxfun(x=tstep,y=c(rep(0,100),rep(80,400)))
st2Fun=approxfun(x=tstep,y=c(rep(4,500))) #this is to simulate increase in fecundity, could change the model to make f time dependent, or have some amount of harvest that declines to 0

simS=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)

#harvesting delays the transition to sp2
tstep=1:500
h1Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,0,length.out = 100), rep(0,300)))
h2Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,10,length.out = 100),rep(10,300)))
qE1Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,4,length.out = 200), rep(4,200)))
qE2Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,.5,length.out = 200), rep(.5,200)))
st1Fun=approxfun(x=tstep,y=c(rep(0,500)))
st2Fun=approxfun(x=tstep,y=c(rep(5,500))) #this is to simulate increase in fecundity, could change the model to make f time dependent, or have some amount of harvest that declines to 0

simH=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)

#harvesting and stocking delays the transition to sp2
tstep=1:500
h1Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,0,length.out = 100), rep(0,300)))
h2Fun=approxfun(x=tstep,y=c(rep(10,100),seq(10,10,length.out = 100),rep(10,300)))
qE1Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,4,length.out = 200), rep(4,200)))
qE2Fun=approxfun(x=tstep,y=c(rep(0,100), seq(0,.5,length.out = 200), rep(.5,200)))
st1Fun=approxfun(x=tstep,y=c(rep(0,100),rep(80,400)))
st2Fun=approxfun(x=tstep,y=c(rep(5,500))) #this is to simulate increase in fecundity, could change the model to make f time dependent, or have some amount of harvest that declines to 0

simB=ode(y=y0,times=tstep,func=simBiggsQ2,parms=p)

#plotting
par(mfcol=c(2,2), mar=c(1,1,1,1), oma=c(4,4,1,1))
plot(sim[,1],sim[,2],type='l',ylim=c(0,max(sim[,2:3],na.rm = T)),col='grey',lwd=2,xlab = "",ylab = "")
lines(sim[,1],sim[,3],col='black',lwd=2)
text(x=375,y=1200,labels = "No action taken")
text(x=max(sim[,1],na.rm = T), y=max(sim[,2:3],na.rm = T),labels='A',font = 2)
legend("topleft", legend = c("sp1","sp2"), col = c('grey','black'),lty = c(1,1),lwd=2,bty = "n")
plot(simS[,1],simS[,2],type='l',ylim=c(0,max(sim[,2:3],na.rm = T)),col='grey',lwd=2,xlab = "",ylab = "")
lines(simS[,1],simS[,3],col='black',lwd=2)
text(x=375, y=1200, labels = "Stocking sp1")
text(x=min(simS[,1],na.rm=T), y=max(simS[,2:3],na.rm = T),labels='B',font = 2)
plot(simH[,1],simH[,2],type='l',ylim=c(0,max(sim[,2:3],na.rm = T)),col='grey',lwd=2,xlab = "",ylab = "")
lines(simH[,1],simH[,3],col='black',lwd=2)
text(x=375, y=1200, labels = "Harvesting sp2")
text(x=min(simH[,1],na.rm = T), y=max(simH[,2:3],na.rm = T),labels='C',font = 2)
plot(simB[,1],simB[,2],type='l',ylim=c(0,max(sim[,2:3],na.rm = T)),col='grey',lwd=2,xlab = "",ylab = "")
lines(simB[,1],simB[,3],col='black',lwd=2)
text(x=375, y=1200, labels = "Harvesting sp2 \n Stock sp1")
text(x=min(simB[,1],na.rm = T), y=max(simB[,2:3],na.rm = T),labels="D",font = 2)
mtext("Year", side = 1, outer = T, line = 2.5)
mtext("Population Size", side = 2, outer = T, line = 2.5)