---
title: "Sensitivity Analysis Doc"
author: "Colin Dassow, Chelsey Neiman, Chris Solomon, Greg Sass, and Stuart Jones"
date: "12/31/2020"
output: word_document
---

# Sensitivity Analysis

Here I systematically vary juvenile survival $s$, adult natural mortality $m$, cannibalism $c_{j_1,a_1}$, predation by adult $c_{j_1,a_2}$, juvenile competition $c_{j_1,j_2}$, and fecundity (Ricker parameters $a$ and $b$) to see how the effect the occurence of stabe states. I did this by only varying values for species 1 and I didn't now vary any of these parms in combination with each other, only in isolation. 

Below are figures to describe the results

```{r setup, include=FALSE}

library(deSolve)
library(ggplot2)
library(ggpubr)
#model function
# qE - harvest, species specific, this is what is controlled by 'regulations'
# s - juvenile overwinter survival, species specific
# m - adult natural mortality rate, species specific
# cJA - effect of adults of a given species on juveniles of a given species (cover cannibalism or interspecific predation, both happen in foraging arena)
# cJJ - effect of juveniles of one species on juveniles of the other (can be predation or competition)
# h - rate at which juveniles leave foraging arena for refuge, species specific
# v - rate at which juveniles enter foraging arena from refuge, species specific
# stock1 - annual stocked num spp1
# stock2 - annual stocked num spp2
simBiggsQ2<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}
```

```{r juvSurvival, echo=FALSE, cash=TRUE, fig.cap="Juvenile Survival rate. Alternate stable states persist for different values, except the lowest value and only when species 2 is initally dominant (0.05). The range of harvest values over which alternate states occur increases as juvenile survival increases. Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

tstep=times=1:300

ss=seq(0.05,.9,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,ss);colnames(combos)=c("qEs","ss")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=store$ss[i],m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
  #store$ss[i]=ss[f]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=store2$ss[i],m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
  #store2$ss[i]=ss[f]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$ss,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$ss,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","ss","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)#;panA$ss=as.numeric(panA$ss)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$ss,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$ss,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","ss","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)#;panB$ss=as.numeric(panB$ss)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=ss))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=ss))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r adultNatMort, echo=FALSE, cash=TRUE, fig.cap="Adult natural mortality rate. Alternate stable states persist for all values. The range of harvest values over which alternate states occur decreases as adult natural mortality decreases. In other words, as fewer adults die annually more harvest can happen on species 1 while still allowing it to dominate (panel A). Panel A - species 2 initially dominant; Panel B - species 1 initially dominant. "}

#looking to see how variation in the adult natural mortality rate for species 1 changes whether or not stable states occur
times=1:300

ms=seq(0.05,.9,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,ms);colnames(combos)=c("qEs","ms")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=store$ms[i],cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=store2$ms[i],cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$ms,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$ms,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","ms","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$ms,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$ms,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","ms","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=ms))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=ms))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r adultPredation, echo=FALSE, cash=TRUE, fig.cap="Species 1 adults predationon species 2 juveniles. Alternate stable states persist for all values except when species 2 is dominant and Adults of species 1 have no effect on species 2 juveniles (0). As the effects of species 1 adults on species 2 juveniles increases the range of harvest values over which alternate stable states occur shifts right (i.e. more harvest can be tolerated before the system flips). Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#looking to see how variation in the adult predation j2 for species 1 changes whether or not stable states occur
times=1:300

cj2a1s=seq(0,1,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,cj2a1s);colnames(combos)=c("qEs","cj2a1s")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=store$cj2a1s[i],cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=store2$cj2a1s[i],cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$cj2a1s,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$cj2a1s,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","cj2a1s","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$cj2a1s,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$cj2a1s,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","cj2a1s","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=cj2a1s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=cj2a1s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))

```

```{r cannibalism, echo=FALSE, cash=TRUE, fig.cap="Species 1 cannibalism rate. Alternate stable states persist for all values. When little cannibalism occurs (0.0001) more harvest can be tolerated, and vice versa when more cannibalism occurs. Most cannibalism rates result in roughly the same flipping points. Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

times=1:300

can=seq(0.0001,.05,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,can);colnames(combos)=c("qEs","can")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)


y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=store$can[i],cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=store2$can[i],cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
  
}

panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$can,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$can,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","can","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$can,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$can,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","can","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=can))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=can))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r competition, echo=FALSE, cash=TRUE, fig.cap="Effect of J1 on J2. Alternate stable states do NOT persist for all values. This parameter has a large effect on if alternate states exist and when the flipping point is crossed. Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#looking to see how variation in the j1 effect on j2 changes whether or not stable states occur
times=1:300

cj2j1s=seq(0,0.5,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,cj2a1s);colnames(combos)=c("qEs","cj2j1s")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=store$cj2j1s[i],v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=store2$cj2j1s[i],v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$cj2j1s,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$cj2j1s,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","cj2j1s","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$cj2j1s,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$cj2j1s,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","cj2j1s","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=cj2j1s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=cj2j1s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r competition2, echo=FALSE, cash=TRUE, fig.cap="Effect of J2 on J1. Alternate stable states do NOT persist for all values. This parameter has a large effect on if alternate states exist and when the flipping point is crossed. Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#looking to see how variation in the j1 effect on j2 changes whether or not stable states occur
times=1:300

cj1j2s=seq(0,0.5,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,cj2a1s);colnames(combos)=c("qEs","cj1j2s")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=store$cj1j2s[i],v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=store$cj1j2s[i],v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$cj1j2s,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$cj1j2s,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","cj1j2s","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$cj1j2s,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$cj1j2s,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","cj1j2s","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=cj1j2s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=cj1j2s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r A2_predation_on_J1, echo=FALSE, cash=TRUE, echo=FALSE, cash=TRUE, fig.cap="Effect of A2 predation on J1. Alternate stable states persist across all values. As the effect of A2 on J1 decreases more harvest of species 1 can be tolerated before the system flips. Larger effect on the upper end of the stable state where the system starts dominated by species 1 and flips to spcies 2 at high harvest (panel B). Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#looking to see how variation in the adult predation j1 for species 2 changes whether or not stable states occur
times=1:300

cj1a2s=seq(0,1,length.out = 5)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,cj2a1s);colnames(combos)=c("qEs","cj1a2s")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=store$cj1a2s[i],cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=store2$cj1a2s[i],cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$cj1a2s,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$cj1a2s,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","cj1a2s","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$cj1a2s,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$cj1a2s,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","cj1a2s","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=cj1a2s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=cj1a2s))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))


```

```{r ricker_a, echo=FALSE, cash=TRUE, fig.cap="Effect varying ricker A parm - max number of recruit produced. Alternate stable states persist across all values except the very lowest a value (1250) and only when species 2 is initially dominant (panel A). Parameter A has a larger effect on when flipping happens in panel B; where species 1 is initially dominant.  Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#### RICKER SENSITIVITY ####
#varying the ricker parms a & b for species 1 to see the effect of differing recruitment on stable states
# I need to create a few functions that have different parms since the function has them coded into it.
# based on Hilborn and Walters book ricker parm a is maybe more important to vary
#species 2 stock recruitment will stay the same through all variants of speces 1 recruitment

#original a=5000; b=500
simBiggsQ2<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

#var1 a=2500; b=500
var1<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(2500*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var2 a=1250; b=500
var2<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(1250*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 3 a=7500; b=500
var3<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(7500*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 4 a=4500; b=500
var4<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(4500*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 5 a=5500; b=500
var5<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5500*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

times=1:300
funlist=list(simBiggsQ2,var1,var2,var3,var4,var5)
as=c(2500,1250,7500,4500,5500,5000)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,as);colnames(combos)=c("qEs","as")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  if(store$as[i]==5000){
    sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  }else{if(store$as[i]==5500){
    sim=ode(y=y0,times=times,func=var5,parms=p)
  }else{if(store$as[i]==4500){
    sim=ode(y=y0,times=times,func=var4,parms=p)
  }else{if(store$as[i]==7500){
    sim=ode(y=y0,times=times,func=var3,parms=p)
  }else{if(store$as[i]==1250){
    sim=ode(y=y0,times=times,func=var2,parms=p)
  }else{if(store$as[i]==2500){
    sim=ode(y=y0,times=times,func=var1,parms=p)
  }
    }}}}}
  
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  if(store2$as[i]==5000){
    sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  }else{if(store2$as[i]==5500){
    sim=ode(y=y0,times=times,func=var5,parms=p)
  }else{if(store2$as[i]==4500){
    sim=ode(y=y0,times=times,func=var4,parms=p)
  }else{if(store2$as[i]==7500){
    sim=ode(y=y0,times=times,func=var3,parms=p)
  }else{if(store2$as[i]==1250){
    sim=ode(y=y0,times=times,func=var2,parms=p)
  }else{if(store2$as[i]==2500){
    sim=ode(y=y0,times=times,func=var1,parms=p)
  }
  }}}}}
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$as,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$as,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","as","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$as,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$as,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","as","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=as))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=as))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))
```

```{r ricker_b, echo=FALSE, cash=TRUE, fig.cap="Effect varying ricker B parm - stock size to produce .5 of A. Alternate stable states persist across all values. Seems to be a large effect of this parameter in either scenario (panel A or B). No general trend here because of the nonlinear effect of ricker parm b on recruitment for a given abundeance. Panel A - species 2 initially dominant; Panel B - species 1 initially dominant."}

#### RICKER PARM B SENSITIVITY ####
#original a=5000; b=500
simBiggsQ2<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(500+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

#var1 a=5000; b=250
var1b<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(250+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var2 a=5000; b=125
var2b<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(1250*A1/(125+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 3 a=5000; b=750
var3b<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(750+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 4 a=5000; b=450
var4b<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(450+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

# var 5 a=5000; b=550
var5b<-function(t,y,params){
  A1<-y[1]
  A2<-y[2]
  J1<-y[3]
  J2<-y[4]
  with(as.list(params),{
    dA1dt=-qE1Fun(t)*A1-m1*A1+s1*J1
    dA2dt=-qE2Fun(t)*A2-m2*A2+s2*J2
    dJ1dt=-cJ1J2*J2*J1-((cJ1A2*v1*A2*J1)/(h1Fun(t)+v1+cJ1A2*A2))-((cJ1A1*v1*A1*J1)/(h1Fun(t)+v1+cJ1A1*A1))-s1*J1+(5000*A1/(550+A1))+st1Fun(t)
    dJ2dt=-cJ2J1*J1*J2-((cJ2A1*v2*A1*J2)/(h2Fun(t)+v2+cJ2A1*A1))-((cJ2A2*v2*A2*J2)/(h2Fun(t)+v2+cJ2A2*A2))-s2*J2+(5000*A2/(500+A2))+st2Fun(t)
    return(list(c(dA1dt,dA2dt,dJ1dt,dJ2dt)))
  })
}

times=1:300
bs=c(250,125,750,450,550,500)
qEs=seq(0,8,length.out=30)
combos=expand.grid(qEs,bs);colnames(combos)=c("qEs","bs")
store=data.frame(A1=0,A2=0,J1=0,J2=0);store=cbind(combos,store)
store2=data.frame(A1=0,A2=0,J1=0,J2=0);store2=cbind(combos,store2)

y0=c(10,100,0,0)
for(i in 1:nrow(store)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(2,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  if(store$bs[i]==500){
    sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  }else{if(store$bs[i]==550){
    sim=ode(y=y0,times=times,func=var5b,parms=p)
  }else{if(store$bs[i]==450){
    sim=ode(y=y0,times=times,func=var4b,parms=p)
  }else{if(store$bs[i]==750){
    sim=ode(y=y0,times=times,func=var3b,parms=p)
  }else{if(store$bs[i]==125){
    sim=ode(y=y0,times=times,func=var2b,parms=p)
  }else{if(store$bs[i]==250){
    sim=ode(y=y0,times=times,func=var1b,parms=p)
  }
  }}}}}
  
  store$A1[i]=sim[nrow(sim)-1,2]
  store$A2[i]=sim[nrow(sim)-1,3]
  store$J1[i]=sim[nrow(sim)-1,4]
  store$J2[i]=sim[nrow(sim)-1,5]
}
y0=c(100,10,40,40)
for(i in 1:nrow(store2)){
  qE1Fun=approxfun(x=tstep,y=c(rep(store2$qEs[i],length(tstep))))
  qE2Fun=approxfun(x=tstep,y=rep(1.8,length(tstep)))
  h1Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  h2Fun=approxfun(x=tstep,y=rep(8,length(tstep)))
  st1Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  st2Fun=approxfun(x=tstep,y=rep(0,length(tstep)))
  
  p=c(c(s1=0.5,m1=0.5,cJ1A1=0.001,cJ1A2=0.5,cJ1J2=0.003,v1=1),
      c(s2=0.5,m2=0.5,cJ2A2=0.001,cJ2A1=0.3,cJ2J1=0.003,v2=1))
  if(store2$bs[i]==500){
    sim=ode(y=y0,times=times,func=simBiggsQ2,parms=p)
  }else{if(store2$bs[i]==550){
    sim=ode(y=y0,times=times,func=var5b,parms=p)
  }else{if(store2$bs[i]==450){
    sim=ode(y=y0,times=times,func=var4b,parms=p)
  }else{if(store2$bs[i]==750){
    sim=ode(y=y0,times=times,func=var3b,parms=p)
  }else{if(store2$bs[i]==125){
    sim=ode(y=y0,times=times,func=var2b,parms=p)
  }else{if(store2$bs[i]==250){
    sim=ode(y=y0,times=times,func=var1b,parms=p)
  }
  }}}}}
  store2$A1[i]=sim[nrow(sim)-1,2]
  store2$A2[i]=sim[nrow(sim)-1,3]
  store2$J1[i]=sim[nrow(sim)-1,4]
  store2$J2[i]=sim[nrow(sim)-1,5]
}


panA=as.data.frame(rbind(cbind(store$qEs,store$A1,store$bs,rep("A1",nrow(store))),cbind(store$qEs,store$A2,store$bs,rep("A2",nrow(store)))))
colnames(panA)=c("qEs","Abund","bs","sp")
panA$qEs=as.numeric(panA$qEs);panA$Abund=as.numeric(panA$Abund)
panB=as.data.frame(rbind(cbind(store2$qEs,store2$A1,store2$bs,rep("A1",nrow(store2))),cbind(store2$qEs,store2$A2,store2$bs,rep("A2",nrow(store2)))))
colnames(panB)=c("qEs","Abund","bs","sp")
panB$qEs=as.numeric(panB$qEs);panB$Abund=as.numeric(panB$Abund)

a=ggplot(data = panA,aes(x=qEs,y=Abund,color=sp,linetype=bs))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
b=ggplot(data = panB,aes(x=qEs,y=Abund,color=sp,linetype=bs))+theme_classic()+
  geom_line(size=1)+scale_color_manual(values = c("black","grey"),name="")+
  scale_linetype_discrete(name="")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
fig1=ggarrange(a,b,ncol=1,labels = c("A","B"), common.legend = T,legend="top",label.x = 0.9)
annotate_figure(fig1,
                left = text_grob("Adult Abundance", rot = 90),
                bottom = text_grob("Harvest Rate (qE)"))

```